Quantifying Stability and Change in Personal Culture Using Panel Data

### 2023-09-25 manuscript.Rmd
Limitations

## Implications for Survey Measurement

@@

The approach outlined here generates estimates of how much variance in a population is due to each of two broad processes: intrapersonal change and interpersonal differences at baseline. It does not quantify the proportion of people who “change.” When we observe intrapersonal change in a question, we cannot be sure whether this is driven by many people making small changes or a few people making large changes. Adjudicating this question will require different models that distinguish different kinds of change.

Our results and interpretations also hinge on how we have defined change and the assumptions built into the Life Course Adaption Model. This model treats cultural trajectories as varying linear slopes for each respondent, assuming that change is a linear function of time. This assumption likely does not hold true in practice. Change over time likely takes non-linear and discontinuous forms. People might jump from one “stable” disposition to another, for example, or experience a discontinuous "turning point" that upsets a stable trajectory. This assumption is in part a limitation of the data, as most questions are only observed for three to six waves with a large number of missing cases. Panels with more waves might allow for more flexible models of change that do not require this assumption of linearity.

Similarly, in assuming a linear trajectory for change, our approach assumes that durable change is unidirectional. To the extent that people make durable change as a result of a social transition and then return to a previous position later in life as a result of a second transition (marriage and divorce, for example), this is captured as residual within-person variance, rather than one of our two broad theoretical processes. This is a potentially reasonable assumption on short panels --change that does not last two years is hard to classify as durable-- but it might not be true in longer panels. Other researchers with more flexible definitions of "durable change" might potentially identify more durable change even in the same population.

Finally, while we examine an incredible broad array of measures of personal culture, from assessments about the relative importance of different features when buying a new car and policy preferences to religious beliefs and core values, we are limited by the kinds of questions that are asked in panel surveys and the contexts in which these surveys were administered (Australia, Germany, Switzerland, the United Kingdom and the United States). This means our findings tend to reflect questions about general (national) politics, gender roles, immigration and race relations, and general well being. Our data does not cover some dimensions of culture that might be relevant for other aspects of sociological exploration (artistic tastes, leisure activities, and time use, for example).

>>>>

## Implications for Survey Research

@@

The approach outlined here generates estimates of how much of the observed variance in a population can be attributed to one of two theoretical processes: intrapersonal change and interpersonal differences at baseline. It does not quantify the proportion of people who “change,” nor can we be sure that the amount of interpersonal change we detect is driven by many people making small changes or a few people making large changes.

Our results and interpretations also hinge on how we have defined change. Similarly to the LCAM, our approach treats cultural trajectories as varying linear slopes for each respondent, thus assuming that change is a linear function of time. There is no doubt, this assumption severely simplifies reality in which change likely also takes non-linear and discontinuous forms. People might jump from one “stable” disposition to another or experience a "turning point" in their lives that upsets an otherwise stable trajectory. In parts, our assumption of linearity is a limitation of the data, as most questions are only observed for three to six waves. Panels with more waves might allow to loosen this assumption to test alternative, more flexible models of change.

Similarly, our approach assumes that durable change is unidirectional. This is a sensible assumption on short panels where classifying change that lasts less than two years as durable seems unreasonable. Practically, it means that people making durable changes in their cultural dispositions to then return to a previous state later in life are classified as neither accounted for by the two theoretical processes. Longer panels and more flexible definitions of "change" would allow us to account for such trajectories and thus potentially also identify more durable change even in the same population.

Finally, we had chosen to examine the broadest array of measures of personal culture available to us, ranging from religious beliefs and core values, to policy preferences, and even to the importance of different features when buying a new car across five countries (Australia, Germany, Switzerland, the United Kingdom and the United States). Nevertheless, our findings remain limited to the kinds of questions that are asked in panel surveys and in the contexts they were administered, reflecting issues of general (national) politics, gender roles, immigration and race relations, and general well being. Although we have no reason to believe results to be different, our findings do not directly speak to other dimensions of culture such as artistic tastes, leisure activities, and time use.
